---
title: "Statistical Inference Project - Part 2"
author: "Andrew Mendonca"
date: "July 28, 2017"
output: pdf_document
---

## Overview
This report will analyze the ToothGrowth data in the R datasets package.

## Analysis
1. Load the data and perform some exploratory data analyses

Load the following library.

```{r, echo = TRUE}
library(ggplot2)
```
Load the data.

```{r, echo = TRUE}
data("ToothGrowth")
```
View all of the dataset variables.

```{r, echo = TRUE}
str(ToothGrowth)
```
Convert the variable dose from numeric to factor
```{r, echo = TRUE}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
```
See the variable after conversion.
```{r, echo = TRUE}
str(ToothGrowth)
```

2. Provide a basic summary of the data.

Summarize all stats for every single variable

```{r, echo = TRUE}
summary(ToothGrowth)
```
Display the first few rows of the ToothGrowth data.
```{r, echo = TRUE}
head(ToothGrowth)
```

3. Use confidence intervals and hypothesis tests to compare tooth growth by supp and dose.

Compare tooth growth by supplement.
```{r, echo = TRUE}
t.test(len ~ supp, data = ToothGrowth)
```
The p-value is 0.06 and the confidence interval is 0. This means that we cannot ignore the
null hypothesis that supplement types have no effect on tooth growth.

Compare tooth growth by dose based on the different pairs of dose values.

```{r, echo = TRUE}
ToothGrowth_subdata <- subset(ToothGrowth, ToothGrowth$dose %in% cbind(1.0, 0.5))
t.test(len ~ dose, data = ToothGrowth_subdata)
ToothGrowth_subdata <- subset(ToothGrowth, ToothGrowth$dose %in% cbind(0.5, 2.0))
t.test(len ~ dose, data = ToothGrowth_subdata)
ToothGrowth_subdata <- subset(ToothGrowth, ToothGrowth$dose %in% cbind(1.0, 2.0))
t.test(len ~ dose, data = ToothGrowth_subdata)
```
The p-value of each test is 0 and the confidence interval of each test is also 0.

So, we can assume that the average tooth length increases with a higher dose, and thus,
the null hypothesis is ignored.

## Conclusions
We made a few assumptions that the sample is representative of the population and the 
distribution of the sample means follows the Central Limit Theorem.

Based on the tests above, we conclude that the supplement delivery method does not have
any effect on tooth growth, but higher dosages result in increased tooth length.