---
title: "Statistical Inference Project - Part 1"
author: "Andrew Mendonca"
date: "July 28, 2017"
output: pdf_document
---

## Overview
In this project you will investigate the exponential distribution in R and compare it with the Central 
Limit Theorem. You will also investigate the distribution of averages of 40 exponentials.

## Simulations
Load the following library for creating graphs.

```{r, echo = TRUE}
library(ggplot2)
```
Set the variables for controlling the simulation.

```{r, echo = TRUE}
set.seed(234)
num_simulation <- 1000
lambda <- 0.2
n <- 40
```
Create a matrix of 1000 simulations each with 40 samples being drawn from the exponential distribution 
and then calculate the averages.

```{r, echo = TRUE}
matrix_simulation <- matrix(rexp(num_simulation * n, rate = lambda), num_simulation, n)
mean_simulation <- rowMeans(matrix_simulation)
```
## Sample Mean versus Theoretical Mean
Calculate both the actual mean for the sample data and the theoretical mean and compare both of them.

```{r, echo = TRUE}
sample_mean <- mean(mean_simulation)
sample_mean
theoretical_mean <- 1 / lambda
theoretical_mean
```
Create a graph to show the comparison between the sample mean and the theoretical mean.

```{r, echo = TRUE}
hist(mean_simulation, main = "Sample Mean vs Theoretical Mean", xlab = "Mean", ylab = "Frequency")
abline(v = sample_mean, col = "orange")
abline(v = theoretical_mean, col = "blue")
```
The actual mean is 5.001573 and the theoretical mean is 5. This shows that the
actual mean from the sample data and the theoretical mean of the normal data
are very close to each other.

## Sample Variance versus Theoretical Variance
Calculate both the actual variance for the sample data and the theoretical variance and compare both of them.

```{r, echo = TRUE}
sample_variance <- var(mean_simulation)
sample_variance
theoretical_variance <- (1 / lambda)^2 / n
theoretical_variance
```
The actual variance is 0.6631504 and the theoretical variance is 0.625. This shows that both variances 
are very close to each other.

Calculate the sample standard of deviation and theoretical standard of deviation.

```{r, echo = TRUE}
sample_stdDev <- sd(mean_simulation)
sample_stdDev
theoretical_stdDev <- 1 / (lambda * sqrt(n))
theoretical_stdDev
```
The sample standard of deviation is 0.8143405 and the theoretical deviation is 0.7905694.

## Distribution
Create a graph to show that the distribution is approximately normal.

```{r, echo = TRUE}
data_display <- data.frame(mean_simulation)
ggplot(data_display, aes(x = mean_simulation)) + geom_histogram(aes(y = ..density..), 
  colour = "black", fill = "orange") + ggtitle("The Distribution of Averages of 40 Exponentials") + 
  xlab("Mean") + ylab("Density") + geom_vline(aes(xintercept = sample_mean, colour = "lightblue")) + 
  geom_vline(aes(xintercept = theoretical_mean, colour = "lightred")) + stat_function(fun = dnorm, 
  args = list(mean = sample_mean, sd = sample_stdDev), color = "green", size = 1.0) + 
  stat_function(fun = dnorm, args = list(mean = theoretical_mean, sd = theoretical_stdDev), 
  color = "violet", size = 1.0)
```
The density of the actual data is shown by orange bars. The sample mean and theoretical mean nearly overlap
each other. The violet line shows the normal curve of the theoretical mean and standard deviation. The
violet line shows the curve of the sample mean and standard deviation. So, the distribution of averages
of 40 exponential distributions is close to the normal distribution.